---
title: "CodingChallenge7"
author: "Olivia Brown"
date: "2025-04-03"
output: word_document
editor_options: 
  chunk_output_type: console
---
[Link to Github](https://github.com/olixiabrown/PLPA5820) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Q1 - Load in required packages and dataset
```{r}
getwd()
plantemergence <- read.csv("CodingChallenge7/PlantEmergence.csv")

library(tidyverse)
library(lme4)
library(emmeans)
library(multcompView) 
library(multcomp)


#turn df vectors into factors
str(plantemergence)
plantemergence$Treatment <- as.factor(plantemergence$Treatment)
plantemergence$DaysAfterPlanting <- as.factor(plantemergence$DaysAfterPlanting)
plantemergence$Rep <- as.factor(plantemergence$Rep)
```

####Q2 - lm with interaction
```{r}
emergence.lm <- lm(Emergence ~ Treatment*DaysAfterPlanting, data = plantemergence)

summary(emergence.lm)
anova(emergence.lm)


```


####Q3 - lm without interaction
We do not need to fit the interaction term because the interactions are not statistically significant, so will use the following model for the rest of the assignment. 
```{r}
emergence.lm2 <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = plantemergence)
summary(emergence.lm2)
anova(emergence.lm2)
```
The estimate of treatment 2 deviates from the intercept (Treatment 1) by 134.531 which is more than the estimates for all other treatments. 

####Q4 - emmeans least squared test
```{r}
lsmeans <- emmeans(emergence.lm2, ~Treatment)
results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE)

```


####Q5
significance of letters: treatments that share a letter are not significantly different from each other and treatments that do not share a letter are significantly different from each other. treatment two is the only treatment that is significantly different from every other treatment. 
```{r}
plot_cldbars_onefactor <- function(emergence.lm2, factor) {
data <- emergence.lm2$model
variables <- colnames(emergence.lm2$model)
dependent_var <- variables[1]
independent_var <- variables[2:length(variables)]

#estimate lsmeans
lsmeans2 <- emmeans(emergence.lm2, as.formula(paste("~", factor))) 
# contrast with Tukey adjustment by default
Results_lsmeans2 <- cld(lsmeans2, alpha = 0.05, reversed = TRUE, details =
TRUE, Letters = letters) 

# Extracting the letters for the bars
sig.diff.letters <- data.frame(Results_lsmeans2$emmeans[,1],
str_trim(Results_lsmeans2$emmeans[,7]))
colnames(sig.diff.letters) <- c(factor, "Letters")


# for plotting with letters from significance test
ave_stand2 <- emergence.lm2$model %>%
group_by(!!sym(factor)) %>%
dplyr::summarize(
ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
se = sd(.data[[dependent_var]]) / sqrt(n())
) %>%
left_join(sig.diff.letters, by = factor) %>%
mutate(letter_position = ave.emerge + 10 * se)


plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) +
stat_summary(fun = mean, geom = "bar") +
stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
ylab("Number of emerged plants") +
geom_jitter(width = 0.02, alpha = 0.5) +
geom_text(data = ave_stand2, aes(label = Letters, y = letter_position),
size = 5) +
xlab(as.character(factor)) +
theme_classic()

return(plot)
}

plot_cldbars_onefactor(emergence.lm2, "Treatment")
```
